---
title: "uiwuev"
author: "uifa"
date: "2024-01-08"
output: html_document
---

```{r}

#Install and Load Required Packages
library(dplyr)
library(ggplot2)
library(openxlsx)
library(ggcorrplot)
library(randomForest)
library(lmtest)
library(AER)

```


```{r}
#Import the data Set
loan_data=read.csv("E:\\3rd Year- UOR\\2nd Sem\\R Programming\\loan.csv",fileEncoding = "latin1")
```


```{r}
#Clean the data
#remove missing values(Null or NA)
loan_data <- na.omit(loan_data)


#Replace Missing Values
#Create a function for replacing missing values with the column mean
replace_na_with_mean <- function(column) {
ifelse(is.na(column), ave(column, FUN = function(x) mean(x, na.rm = TRUE)), column)
  
# Remove duplicate rows
loan_data <- loan_data[!duplicated(loan_data), ]


  
  
}

#List of columns to process
columns_to_process <- c("loan_amnt", "term", "int_rate", "installment", 
                         "grade", "emp_length", "home_ownership", 
                         "annual_inc", "loan_status", "purpose", 
                         "dti", "delinq_2yrs", "open_acc", 
                         "pub_rec", "revol_util", "total_acc", 
                         "total_pymnt_inv", "total_rec_prncp", 
                         "total_rec_int", "repay_fail")
# Apply the function to each column in the list
loan_data[, columns_to_process] <- lapply(loan_data[, columns_to_process], replace_na_with_mean)


#Covert categorical variables values into Numerical variables values
loan_data[c("term", "grade", "emp_length", "home_ownership", "loan_status", "purpose")] <- lapply(loan_data[c("term", "grade", "emp_length", "home_ownership", "loan_status", "purpose")], factor)
loan_data <- na.omit(loan_data)

#Write the new excel file 
write.csv(loan_data, "output_file.csv", row.names = FALSE)

#Import new data
loan_data_new=read.csv("C:\\Users\\Supun\\Desktop\\Final\\output_file.csv")

#Explore the structure of the dataset
loan_data_new <- na.omit(loan_data_new)
View(loan_data_new)
```



```{r}
#Calculate the correlation matrix
cor_matrix <- cor(loan_data_new[, sapply(loan_data_new, is.numeric)])
#Visualizing the correlation matrix with ggcorrplot
ggcorrplot(cor_matrix,hc.order = TRUE,type="lower",lab = TRUE,lab_size = 1)
#Building a random forest model to predict 'repay_fail' based on all features
model_rf <- randomForest(repay_fail ~ ., data = loan_data_new)
varImpPlot(model_rf)
```

01) logit:

Description: The logistic link function is commonly used for binary regression , predictors to the range [0, 1], representing probabilities.

```{r}
#model of the logit link function
model_logit<- glm(repay_fail ~ installment + total_pymnt_inv + loan_amnt + total_rec_prncp + loan_status, 
            family = binomial(link = "logit"), data = loan_data_new)

plot(model_logit)
summary(model_logit)
```

02) probit:

Description: The probit link function is an alternative to the logistic link and is also used for binary regression. It is based on the cumulative distribution function (CDF) of the standard normal distribution.

```{r}
#model of the probit link function
model_probit <- glm(repay_fail ~ installment + total_pymnt_inv + loan_amnt + total_rec_prncp + loan_status, 
            family = binomial(link = "probit"), data = loan_data_new)

plot(model_probit)
summary(model_probit)
```


03) cloglog:

Description: The complementary log-log link function is often used for survival analysis and binary regression.

```{r}
#model of the cloglog link function
model_cloglog <- glm(repay_fail ~ installment + total_pymnt_inv + loan_amnt + total_rec_prncp + loan_status, 
            family = binomial(link = "cloglog"), data = loan_data_new)

plot(model_cloglog)
summary(model_cloglog)
```


```{r}
#Assuming 'model_logit' is your logistic regression model

#1.inearity of Log Odds
#Diagnostic Plot
plot(model_logit$fitted.values, residuals(model_logit, type = "pearson"),
     xlab = "Fitted values", ylab = "Pearson Residuals")

#2.Independence of Observations
#Durbin-Watson Test
dwtest(model_logit)

#3.Homoscedasticity
#Scale-Location Plot
plot(predict(model_logit), sqrt(abs(residuals(model_logit))),
     xlab = "Fitted values", ylab = "Square root of absolute Pearson Residuals")

#4.Deviance Residuals
#Residual vs. Fitted Values Plot
plot(fitted(model_logit), residuals(model_logit, type = "deviance"),
     xlab = "Fitted values", ylab = "Deviance Residuals")

#5.Overdispersion
#Dispersion Test
bptest(model_logit)

#Additional checks as needed for your specific model

#For model summary and overall assessment
summary(model_logit)
```


Comparing AIC values of each model :
model_logit AIC:1530.4  < model_probit AIC: 1602.3 < model_cloglog AIC: 1615.1

using these 3 models,model_logit has the minimum AIC value so we use model_Logit link function to get prediction. 


```{r}
#The predicted values represent the estimated probabilities of success
new_data <- data.frame(installment = c(240,457.25), total_pymnt_inv = c(7811.26,5368.15), loan_amnt = c(9000,6250), total_rec_prncp  = c(4500.00,7999.97), loan_status = c(2,1))
#plot(loan_data_new$repay_fail, predict(model_logit, type = "response"), main = "Actual vs. Predicted", xlab = "Actual", ylab = "Predicted")

predictions <- predict(model_logit, newdata = new_data, type = "response")
predictions
```

